---
title: "Example Graphs from Tostan Data Input"
output: html_notebook
---
```{r, echo=FALSE}
# The following is some preparation and formatting code
# load needed libraries
library(RColorBrewer)
library(lattice) 
library(dplyr)

# Some graph color settings, can ignore for now.
myColours <- brewer.pal(4,"Blues")

my.settings <- list(
  superpose.polygon=list(col=myColours[2:4], border="transparent"),
  strip.background=list(col=myColours[1]),
  strip.border=list(col="black")
)
```

First, import all of the data from the original .csv files. We are also giving every country .csv file a country name and whether the data is EF or EB.

```{r}
source("process_country.R")

Mali <- process_country("17 02 21 MALI_EF CG3_Questionnaire individuel-v2.csv","Mali","EF")
Mauri <- process_country("17 02 21 MAUR_EF CG3_Questionnaire individuel-v2.csv","Mauritania","EF")
GB <- process_country("17-02-21-GB_EF_CG3_Questionnaire individuel-v2.csv","Guinea_Bisseau","EF")

# Take each country and put them together into one large dataset called total
total <- rbind(Mali, Mauri,GB)
```
Next, do some filtering/grouping/calculations to make the data into a useful format. Notice how here the commands are ordered to

(1) Filter the original data, saving the result as 'filtered'

(2) Group the filtered data, saving the result as 'grouped'

(3) Calculate the desired statistics from the grouped data, saving the result as 'result'
```{r}
# Number registered for elections

filtered <- filter(total,village_type=='beneficiary' | is.na(village_type), interviewer!='membre_simple')
grouped <- group_by(filtered, country, interviewer, period)
result <- summarize(grouped,
                    n = n(),
                    num_registered = sum(electoral_list_inscription=='oui', na.rm=TRUE),
                    num_total = length(!is.na(electoral_list_inscription)),
                    percent_registered = num_registered / num_total
)
result 

barchart(percent_registered*100~period|country,
         groups=interviewer,
         data=result,
         ylim=c(0,100),
         ylab = "Percentage Registered",
         main = "Percent Registered to Vote by Country",
        auto.key=list(space="bottom", columns=3, title="Member", cex.title=1),
        par.settings = my.settings
        )
```
```{r}
# Number of children declared. Not getting same n as reported in the result graphs given to me.
total$declared_all_children <- (total$child_under_one_declared + total$child_between_1_and_5_declared) == total$nombre_denfants
filtered <- filter(total, village_type=='beneficiary', nombre_denfants>0) 
grouped <- group_by(filtered, country, interviewer, period)
result <- summarize(grouped,
                    n = n(),
                    number_with_all_declared = length(declared_all_children[declared_all_children==TRUE]),
                    percent_declared = number_with_all_declared / n
)
result 

barchart(percent_declared*100~period|country,
         groups=interviewer,
         data=result,
         ylim=c(0,100),
         ylab = "Percentage Declared",
         main = "Percent Declared All Children",
        auto.key=list(space="bottom", columns=3, title="Member Type", cex.title=1),
        par.settings = my.settings
        )
```
